---
title: "EDA"
author: "Stuart Miller"
date: "November 4, 2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, setup, warning=FALSE, message=FALSE, include=FALSE}

library(tidyverse)
library(gridExtra)
library(rebus)
library(epitools)
library(car)

mysummary<-function(x){
  result<-c(length(x),mean(x),sd(x),sd(x)/length(x),min(x),max(x),IQR(x))
  names(result)<-c("N","Mean","SD","SE","Min","Max","IQR")
  return(result)
}

# load data
data <- 
  as.data.frame(
    read_csv('../diabetic_data.csv',
             col_types = cols(
               encounter_id = col_integer(),
               patient_nbr = col_integer(),
               race = col_factor(levels = c("?", "AfricanAmerican", "Asian", "Caucasian", "Hispanic", "Other")),
               gender = col_factor(levels = c("Male","Female")),
               age = col_factor(levels = c("[0-10)", "[10-20)", "[20-30)","[30-40)","[40-50)","[50-60)",
                                           "[60-70)","[70-80)","[80-90)","[90-100)")),
               weight = col_character(),
               admission_type_id = col_integer(),
               discharge_disposition_id = col_integer(),
               admission_source_id = col_integer(),
               time_in_hospital = col_integer(),
               payer_code = col_character()
             )
    )
  )

```

## EDA

### Some Cleaning

Some cleaning from Dan's research

```{r}
# NA coersion
data <- data %>% mutate_all(~na_if(., '?'))
# Remove observations where discharge dispostion is related to hospice or death
data <- data %>% filter(!data$discharge_disposition_id %in% c('11','13','14','19','20','21'))
# Let's only use the first encounter for a given patient_nbr. Assume the first encounter is the lowest.
data <- data %>% group_by(patient_nbr) %>% filter(encounter_id == min(encounter_id))
# recode the outcomes to yes/no
data$outcome <- factor(ifelse(data$readmitted == "<30","Yes","No"))
```


### Demographics

Demographic variables

* race
* gender
* age
* weight

#### Race

Most patients are Caucasian followed by African American.

```{r}
p1 <- data %>%
  ggplot(aes(x = race)) +
  geom_bar()
p2 <- data %>%
  ggplot(aes(x = race, fill = outcome)) +
  geom_bar(position = 'fill') +
  coord_flip()
grid.arrange(p1, p2, ncol = 2)
```

#### Gender

Gender is close to an even split of the patient records. 

```{r, fig.width=12}
p1 <- data %>%
  ggplot(aes(x = gender)) +
  geom_bar()
p2 <- data %>%
  ggplot(aes(x = gender, fill = outcome)) +
  geom_bar(position = 'fill') +
  coord_flip()
grid.arrange(p1, p2, ncol = 2)
```

#### Age

* The bulk of the patients are in the age range of 40 - 90 years.
* The proportion of readimts increase from the lowest bucket up to the 80s bucket.

```{r, fig.width=12}
p1 <- data %>%
  ggplot(aes(x = age)) +
  geom_bar()
p2 <- data %>%
  ggplot(aes(x = age, fill = outcome)) +
  geom_bar(position = 'fill') +
  coord_flip()
grid.arrange(p1, p2, ncol = 2)
```

#### Weight

Weight is almost entirely missing from the patient records.

```{r}
data %>%
  ggplot(aes(x = weight)) +
  geom_bar()
```


### Hospital Infromation

Hospital Information Variables

* Admission Type ID
* Discharge Disposition ID
* Admission Source ID
* Time in Hospital
* Payer Code
* Medical Specialty

#### Admission Type ID

The levels of this factor are 1-8.
The bulk of the data come from type 1, followed by 2 and 3, then 5 and 6.
Finally, a small amount from 8.

```{r}
p1 <- data %>%
  ggplot(aes(x = admission_type_id)) +
  geom_bar()
p2 <- data %>%
  ggplot(aes(x = admission_type_id, fill = outcome)) +
  geom_bar(position = 'fill') +
  coord_flip()
grid.arrange(p1, p2, ncol = 2)
```

#### Discharge Disposition ID

The levels of this factor are 1-29. 
The bulk of the data come from type 1.
The next highest levels are types 3 and 6.

```{r}
p1 <- data %>%
  ggplot(aes(x = discharge_disposition_id)) +
  geom_bar()
p2 <- data %>%
  ggplot(aes(x = discharge_disposition_id, fill = outcome)) +
  geom_bar(position = 'fill') +
  coord_flip()
grid.arrange(p1, p2, ncol = 2)
```

#### Admission Source ID

This is a 21 level factor variable.
Most obervations are type 6, followed by type 1.

```{r}
p1 <- data %>%
  ggplot(aes(x = admission_source_id)) +
  geom_bar()
p2 <- data %>%
  ggplot(aes(x = admission_source_id, fill = outcome)) +
  geom_bar(position = 'fill') +
  coord_flip()
grid.arrange(p1, p2, ncol = 2)
```

#### Time in Hospital

This is the number of days a patient spent in the hospital.
The distribution is at max at 3 days, then tails off at the time increases.
Maximum observed value is 14 days.

```{r}
p1 <- data %>%
  ggplot(aes(x = time_in_hospital)) +
  geom_bar()
p2 <- data %>%
  ggplot(aes(x = time_in_hospital, fill = outcome)) +
  geom_bar(position = 'fill') +
  coord_flip()
p3 <- data %>%
  ggplot(aes(x = outcome, y = time_in_hospital)) +
  geom_boxplot()
grid.arrange(p1, p2, p3, ncol = 2)
```

#### Payer Code

Many of the payer codes are unknown.
The most common known payer code is `MC`.
Most options have relatively low frequencies.

```{r}
p1 <- data %>%
  ggplot(aes(x = payer_code)) +
  geom_bar()
p2 <- data %>%
  ggplot(aes(x = payer_code, fill = outcome)) +
  geom_bar(position = 'fill') +
  coord_flip()
grid.arrange(p1, p2, ncol = 2)
```


#### Medical Specialty

Most records list unknown medical speciality.
This could be due to the hospital not having a medical speciality.

```{r}
data %>%
  ggplot(aes(x = medical_specialty)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle = 90))
```


### Encounter Information

* Number of Lab Procedures
* Number of Procedures
* Number of Medications
* Number of Outpatient Visits
* Number of Inpatient Visits
* Number of Emergecy Visits

#### Number of Lab Procedures

The number of lab procedures appears to be consistent across the response classes.

```{r}
p1 <- data %>%
  ggplot(aes(x = num_lab_procedures)) +
  geom_histogram(bins = 30) 
p2 <- data %>%
  ggplot(aes(y = num_lab_procedures)) +
  geom_boxplot() + 
  facet_wrap(. ~ readmitted)
grid.arrange(p1,p2, ncol = 2)
```


#### Number of Procedures

The number of procedures appears to be consistent across the response classes.

```{r}
p1 <- data %>%
  ggplot(aes(x = num_procedures)) +
  geom_bar() 
p2 <- data %>%
  ggplot(aes(y = num_procedures)) +
  geom_boxplot() + 
  facet_wrap(. ~ readmitted)
grid.arrange(p1,p2, ncol = 2)
```


#### Number of Medications

The number of medications appears to be consistent across the response classes.

```{r}
p1 <- data %>%
  ggplot(aes(x = num_medications)) +
  geom_histogram(bins = 30) 
p2 <- data %>%
  ggplot(aes(y = num_medications)) +
  geom_boxplot() + 
  facet_wrap(. ~ readmitted)
grid.arrange(p1,p2, ncol = 2)
```


#### Number of Outpatient Visits

The number of recorded outpatient visits is dominated by 0.

```{r}
p1 <- data %>%
  ggplot(aes(x = number_outpatient)) +
  geom_histogram(bins = 30) 
p2 <- data %>%
  ggplot(aes(y = number_outpatient)) +
  geom_boxplot() + 
  facet_wrap(. ~ readmitted)
grid.arrange(p1,p2, ncol = 2)
```


#### Number of Inpatient Visits

The number of recorded inpatient visits is dominated by 0.
There does appear to be a difference between the distributes of inpatient visits and readmittence rate.

```{r}
p1 <- data %>%
  ggplot(aes(x = number_inpatient)) +
  geom_histogram(bins = 30) 
p2 <- data %>%
  ggplot(aes(y = number_inpatient)) +
  geom_boxplot() + 
  facet_wrap(. ~ readmitted)
grid.arrange(p1,p2, ncol = 2)
```


#### Number of Emergency Visits

The number of recorded emergency visits is dominated by 0.

```{r}
p1 <- data %>%
  ggplot(aes(x = number_emergency)) +
  geom_histogram(bins = 30) 
p2 <- data %>%
  ggplot(aes(y = number_emergency)) +
  geom_boxplot() + 
  facet_wrap(. ~ readmitted)
grid.arrange(p1,p2, ncol = 2)
```


### Diagnosis

* Diagnosis 1
* Diagnosis 2
* Diagnosis 3
* Number of Diagnoses

#### Diagnosis 1

Very large number of factor levels, just displaying the top ones here.

```{r}
data$diag_1 <- as.factor(data$diag_1)

data %>%
  mutate(diag_1 = as.factor(diag_1)) %>%
  group_by(diag_1) %>%
  summarise(cnt = n()) %>%
  filter(cnt > 2000) %>%
  ggplot(aes(x = reorder(diag_1, -cnt), y= cnt)) +
  geom_bar(stat="identity") 
```


#### Diagnosis 2

Very large number of factor levels, just displaying the top ones here.

```{r}
data$diag_2 <- as.factor(data$diag_2)

data %>%
  mutate(diag_2 = as.factor(diag_2)) %>%
  group_by(diag_2) %>%
  summarise(cnt = n()) %>%
  filter(cnt > 2000) %>%
  ggplot(aes(x = reorder(diag_2, -cnt), y= cnt)) +
  geom_bar(stat="identity") 
```


#### Diagnosis 3

Very large number of factor levels, just displaying the top ones here.

```{r}
data$diag_3 <- as.factor(data$diag_3)

data %>%
  mutate(diag_3 = as.factor(diag_3)) %>%
  group_by(diag_3) %>%
  summarise(cnt = n()) %>%
  filter(cnt > 1500) %>%
  ggplot(aes(x = reorder(diag_3, -cnt), y= cnt)) +
  geom_bar(stat="identity") 
```


#### Number of Diagnoses

```{r}
p1 <- data %>%
  ggplot(aes(x = number_diagnoses)) +
  geom_histogram(bins = 30) 
p2 <- data %>%
  ggplot(aes(y = number_diagnoses)) +
  geom_boxplot() +
  facet_wrap(. ~ outcome)
p3 <- data %>%
  ggplot(aes(x = number_diagnoses, fill = outcome)) +
  geom_bar(position = 'fill') +
  coord_flip()
grid.arrange(p1, p2, p3, ncol = 2)
```

### Lab Results and Medications

* Glucose Serum Test
* A1c Test Result
* Change of Medications
* Diabetes Medications
* 24 features for medications

#### Glucose Serum Test

```{r}
p1 <- data %>%
  ggplot(aes(x = max_glu_serum)) +
  geom_bar()
p2 <- data %>%
  ggplot(aes(x = max_glu_serum, fill = outcome)) +
  geom_bar(position = 'fill') +
  coord_flip()
grid.arrange(p1, p2, ncol = 2)
```

#### A1c Test Result

```{r}
p1 <- data %>%
  ggplot(aes(x = A1Cresult)) +
  geom_bar()
p2 <- data %>%
  ggplot(aes(x = A1Cresult, fill = outcome)) +
  geom_bar(position = 'fill') +
  coord_flip()
grid.arrange(p1, p2, ncol = 2)
```


#### Change of Medications

```{r}
p1 <- data %>%
  ggplot(aes(x = change)) +
  geom_bar()
p2 <- data %>%
  ggplot(aes(x = change, fill = outcome)) +
  geom_bar(position = 'fill') +
  coord_flip()
grid.arrange(p1, p2, ncol = 2)
```

#### Diabetes Medications

```{r}
p1 <- data %>%
  ggplot(aes(x = diabetesMed)) +
  geom_bar()
p2 <- data %>%
  ggplot(aes(x = diabetesMed, fill = outcome)) +
  geom_bar(position = 'fill') +
  coord_flip()
grid.arrange(p1, p2, ncol = 2)
```


#### Metformin

```{r}
p1 <- data %>%
  ggplot(aes(x = metformin)) +
  geom_bar()
p2 <- data %>%
  ggplot(aes(x = metformin, fill = outcome)) +
  geom_bar(position = 'fill') +
  coord_flip()
grid.arrange(p1, p2, ncol = 2)
```

#### Repaglinide

```{r}
p1 <- data %>%
  ggplot(aes(x = repaglinide)) +
  geom_bar()
p2 <- data %>%
  ggplot(aes(x = repaglinide, fill = outcome)) +
  geom_bar(position = 'fill') +
  coord_flip()
grid.arrange(p1, p2, ncol = 2)
```

#### Nateglinide

```{r}
p1 <- data %>%
  ggplot(aes(x = nateglinide)) +
  geom_bar()
p2 <- data %>%
  ggplot(aes(x = nateglinide, fill = outcome)) +
  geom_bar(position = 'fill') +
  coord_flip()
grid.arrange(p1, p2, ncol = 2)
```

#### Chlorpropamide

```{r}
p1 <- data %>%
  ggplot(aes(x = chlorpropamide)) +
  geom_bar()
p2 <- data %>%
  ggplot(aes(x = chlorpropamide, fill = outcome)) +
  geom_bar(position = 'fill') +
  coord_flip()
grid.arrange(p1, p2, ncol = 2)
```

#### Glimepiride

```{r}
p1 <- data %>%
  ggplot(aes(x = glimepiride)) +
  geom_bar()
p2 <- data %>%
  ggplot(aes(x = glimepiride, fill = outcome)) +
  geom_bar(position = 'fill') +
  coord_flip()
grid.arrange(p1, p2, ncol = 2)
```

#### Acetohexamide

```{r}
p1 <- data %>%
  ggplot(aes(x = acetohexamide)) +
  geom_bar()
p2 <- data %>%
  ggplot(aes(x = acetohexamide, fill = outcome)) +
  geom_bar(position = 'fill') +
  coord_flip()
grid.arrange(p1, p2, ncol = 2)
```

#### Glipizide

```{r}
p1 <- data %>%
  ggplot(aes(x = glipizide)) +
  geom_bar()
p2 <- data %>%
  ggplot(aes(x = glipizide, fill = outcome)) +
  geom_bar(position = 'fill') +
  coord_flip()
grid.arrange(p1, p2, ncol = 2)
```

#### Glyburide

```{r}
p1 <- data %>%
  ggplot(aes(x = glyburide)) +
  geom_bar()
p2 <- data %>%
  ggplot(aes(x = glyburide, fill = outcome)) +
  geom_bar(position = 'fill') +
  coord_flip()
grid.arrange(p1, p2, ncol = 2)
```

#### Tolbutamide

```{r}
p1 <- data %>%
  ggplot(aes(x = tolbutamide)) +
  geom_bar()
p2 <- data %>%
  ggplot(aes(x = tolbutamide, fill = outcome)) +
  geom_bar(position = 'fill') +
  coord_flip()
grid.arrange(p1, p2, ncol = 2)
```

#### Pioglitazone

```{r}
p1 <- data %>%
  ggplot(aes(x = pioglitazone)) +
  geom_bar()
p2 <- data %>%
  ggplot(aes(x = pioglitazone, fill = outcome)) +
  geom_bar(position = 'fill') +
  coord_flip()
grid.arrange(p1, p2, ncol = 2)
```

#### Rosiglitazone

```{r}
p1 <- data %>%
  ggplot(aes(x = rosiglitazone)) +
  geom_bar()
p2 <- data %>%
  ggplot(aes(x = rosiglitazone, fill = outcome)) +
  geom_bar(position = 'fill') +
  coord_flip()
grid.arrange(p1, p2, ncol = 2)
```

#### Acarbose

```{r}
p1 <- data %>%
  ggplot(aes(x = acarbose)) +
  geom_bar()
p2 <- data %>%
  ggplot(aes(x = acarbose, fill = outcome)) +
  geom_bar(position = 'fill') +
  coord_flip()
grid.arrange(p1, p2, ncol = 2)
```

#### Miglitol

```{r}
p1 <- data %>%
  ggplot(aes(x = miglitol)) +
  geom_bar()
p2 <- data %>%
  ggplot(aes(x = miglitol, fill = outcome)) +
  geom_bar(position = 'fill') +
  coord_flip()
grid.arrange(p1, p2, ncol = 2)
```

#### Troglitazone

```{r}
p1 <- data %>%
  ggplot(aes(x = troglitazone)) +
  geom_bar()
p2 <- data %>%
  ggplot(aes(x = troglitazone, fill = outcome)) +
  geom_bar(position = 'fill') +
  coord_flip()
grid.arrange(p1, p2, ncol = 2)
```

#### Tolazamide

```{r}
p1 <- data %>%
  ggplot(aes(x = tolazamide)) +
  geom_bar()
p2 <- data %>%
  ggplot(aes(x = tolazamide, fill = outcome)) +
  geom_bar(position = 'fill') +
  coord_flip()
grid.arrange(p1, p2, ncol = 2)
```

#### Examide

```{r}
p1 <- data %>%
  ggplot(aes(x = examide)) +
  geom_bar()
p2 <- data %>%
  ggplot(aes(x = examide, fill = outcome)) +
  geom_bar(position = 'fill') +
  coord_flip()
grid.arrange(p1, p2, ncol = 2)
```

#### Citoglipton

```{r}
p1 <- data %>%
  ggplot(aes(x = citoglipton)) +
  geom_bar()
p2 <- data %>%
  ggplot(aes(x = citoglipton, fill = outcome)) +
  geom_bar(position = 'fill') +
  coord_flip()
grid.arrange(p1, p2, ncol = 2)
```

#### Insulin

```{r}
p1 <- data %>%
  ggplot(aes(x = citoglipton)) +
  geom_bar()
p2 <- data %>%
  ggplot(aes(x = citoglipton, fill = outcome)) +
  geom_bar(position = 'fill') +
  coord_flip()
grid.arrange(p1, p2, ncol = 2)
```

#### Glyburide-metformin

```{r}
p1 <- data %>%
  ggplot(aes(x = `glyburide-metformin`)) +
  geom_bar()
p2 <- data %>%
  ggplot(aes(x = `glyburide-metformin`, fill = outcome)) +
  geom_bar(position = 'fill') +
  coord_flip()
grid.arrange(p1, p2, ncol = 2)
```

#### Glipizide-metformin

```{r}
p1 <- data %>%
  ggplot(aes(x = `glipizide-metformin`)) +
  geom_bar()
p2 <- data %>%
  ggplot(aes(x = `glipizide-metformin`, fill = outcome)) +
  geom_bar(position = 'fill') +
  coord_flip()
grid.arrange(p1, p2, ncol = 2)
```

#### Glimizide-pioglitazone

```{r}
p1 <- data %>%
  ggplot(aes(x = `glimepiride-pioglitazone`)) +
  geom_bar()
p2 <- data %>%
  ggplot(aes(x = `glimepiride-pioglitazone`, fill = outcome)) +
  geom_bar(position = 'fill') +
  coord_flip()
grid.arrange(p1, p2, ncol = 2)
```

#### Metformin-rosiglitazone

```{r}
p1 <- data %>%
  ggplot(aes(x = `metformin-rosiglitazone`)) +
  geom_bar()
p2 <- data %>%
  ggplot(aes(x = `metformin-rosiglitazone`, fill = outcome)) +
  geom_bar(position = 'fill') +
  coord_flip()
grid.arrange(p1, p2, ncol = 2)
```

#### Metformin-pioglitazone

```{r}
p1 <- data %>%
  ggplot(aes(x = `metformin-pioglitazone`)) +
  geom_bar()
p2 <- data %>%
  ggplot(aes(x = `metformin-pioglitazone`, fill = outcome)) +
  geom_bar(position = 'fill') +
  coord_flip()
grid.arrange(p1, p2, ncol = 2)
```


### Interactions

```{r}
table(data[,c(5,10,51)])

mantelhaen.test(table(data[,c(5,10,51)]))
```


### Intraction Profile Plots

#### Age and Time in Hospital

There is separation in the mean time in hospital for each age group,
but the overlap of varaince is large.

```{r}
sumstats<-aggregate(time_in_hospital~outcome*age,data=data,mysummary)
sumstats<-cbind(sumstats[,1:2],sumstats[,-(1:2)])

ggplot(sumstats,aes(x=age,y=Mean,group=outcome,color=outcome)) +
  ylab("Time in Hospital") +
  xlab("Patient Age") +
  geom_line() +
  geom_point() +
  geom_errorbar(aes(ymin=Mean-SD,ymax=Mean+SD),width=.1)
```

#### Age and Number of Diagnoses

```{r}
sumstats<-aggregate(number_diagnoses~outcome*age,data=data,mysummary)
sumstats<-cbind(sumstats[,1:2],sumstats[,-(1:2)])

ggplot(sumstats,aes(x=age,y=Mean,group=outcome,color=outcome)) +
  ylab("Number of Diagnoses") +
  xlab("Patient Age") +
  geom_line() +
  geom_point() +
  geom_errorbar(aes(ymin=Mean-SD,ymax=Mean+SD),width=.1)
```



#### Discharge Disposition ID and Time in Hospital


```{r}
sumstats<-aggregate(time_in_hospital~outcome*discharge_disposition_id,data=data,mysummary)
sumstats<-cbind(sumstats[,1:2],sumstats[,-(1:2)])

ggplot(sumstats,aes(x=discharge_disposition_id,y=Mean,group=outcome,color=outcome)) +
  ylab("Time in Hospital") +
  xlab("Discharge Disposition ID") +
  geom_line() +
  geom_point() +
  geom_errorbar(aes(ymin=Mean-SD,ymax=Mean+SD),width=.1)
```

#### Discharge Disposition ID and Number of Diagnoses

```{r}
sumstats<-aggregate(number_diagnoses~outcome*discharge_disposition_id,data=data,mysummary)
sumstats<-cbind(sumstats[,1:2],sumstats[,-(1:2)])

ggplot(sumstats,aes(x=discharge_disposition_id,y=Mean,group=outcome,color=outcome)) +
  ylab("Number of Diagnoses") +
  xlab("Discharge Disposition ID") +
  geom_line() +
  geom_point() +
  geom_errorbar(aes(ymin=Mean-SD,ymax=Mean+SD),width=.1)
```

#### 


```{r}
data %>%
  ggplot(aes(x = time_in_hospital, y = number_diagnoses, color = outcome)) +
  geom_jitter(position = "jitter", alpha = 0.1) 
```


####

```{r}
readmit <- 
  data %>%
  filter(outcome %in% 'Yes') %>%
  group_by(age) %>%
  summarise(cnt = n())

tot.per.age <- 
  data %>%
  group_by(age) %>%
  summarise(cnt = n())

readmit.prob.per.age <- readmit[, 2]/tot.per.age[, 2]

logit(readmit.prob.per.age) + 2
```




```{r}
library(pwr2)
pwr.2way(a=2, b=2, alpha=0.05, size.A=30, size.B=30, f.A=NULL, f.B=NULL,
delta.A=4, delta.B=2, sigma.A=2, sigma.B=2)
```


































